\documentclass[aspectratio=169]{beamer}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tcolorbox}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{xcolor}

% Custom colors
\definecolor{primary}{RGB}{41, 128, 185}
\definecolor{secondary}{RGB}{52, 152, 219}
\definecolor{accent}{RGB}{231, 76, 60}
\definecolor{lightgray}{RGB}{236, 240, 241}

% Theme customization
\usetheme{Madrid}
\usecolortheme{whale}
\setbeamercolor{structure}{fg=primary}
\setbeamercolor{background canvas}{bg=white}
\setbeamercolor{normal text}{fg=black}

% Title page info
\title{Pre-Calculus 11}
\subtitle{Solving Quadratic Functions by Factoring}
\author{Created by Yi-Chen Lin}
\date{\today}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

% I) WHAT DOES IT MEAN TO SOLVE AN EQUATION?
\begin{frame}{What Does "Solving" Mean?}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Definition]
        \footnotesize
        \begin{itemize}
            \item Finding a value for "x" (Variable) so that both sides of an equation will be equal
            \item \textbf{Example:} Solve for "x"
            \begin{align*}
                2x + 1 &= 5 \\
                2x &= 4 \\
                x &= 2
            \end{align*}
            \item \textbf{Check:}
            \begin{align*}
                2(2) + 1 &= 5 \\
                4 + 1 &= 5 \\
                5 &= 5 \quad \checkmark
            \end{align*}
        \end{itemize}
    \end{tcolorbox}
\end{frame}

% II) SOLVING TRINOMIALS BY FACTORING
\begin{frame}{Solving Trinomials by Factoring}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Key Concept]
        \footnotesize
        \begin{itemize}
            \item When you have the product of two brackets equal to zero, you can solve this equation easily
            \item \textbf{Example:} Solve $(x + 3)(2x - 5) = 0$
            \begin{align*}
                x + 3 &= 0 \quad \text{or} \quad 2x - 5 = 0 \\
                x &= -3 \quad \text{or} \quad x = 2.5
            \end{align*}
            \item \textcolor{accent}{\textbf{Important Rule:}} Zero times anything is always equal to zero
        \end{itemize}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Practice: Solving Trinomials}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Solve Each of the Following]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item $(x + 4)(x - 3) = 0$
            \item $(2x + 5)(3x - 2) = 0$
            \item $(x - 6)(x + 2) = 0$
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Solutions: Solving Trinomials}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solutions]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item $(x + 4)(x - 3) = 0$
            \begin{align*}
                x + 4 &= 0 \quad \text{or} \quad x - 3 = 0 \\
                x &= -4 \quad \text{or} \quad x = 3
            \end{align*}
            
            \item $(2x + 5)(3x - 2) = 0$
            \begin{align*}
                2x + 5 &= 0 \quad \text{or} \quad 3x - 2 = 0 \\
                x &= -\frac{5}{2} \quad \text{or} \quad x = \frac{2}{3}
            \end{align*}
            
            \item $(x - 6)(x + 2) = 0$
            \begin{align*}
                x - 6 &= 0 \quad \text{or} \quad x + 2 = 0 \\
                x &= 6 \quad \text{or} \quad x = -2
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

% III) ADVANCED FACTORING AND SOLVING
\begin{frame}{Steps for Solving Quadratic Equations}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Key Steps]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Move all terms to one side and make it equal to zero
            \item Factor the equation into two binomials
            \item Make each binomial equal to zero
            \item Solve for "x" from each bracket
        \end{enumerate}
        
        \begin{example}
            Solve: $x^2 + 10x = 9$
            \begin{align*}
                x^2 + 10x - 9 &= 0 \\
                (x + 3)(x - 3) &= 0 \\
                x + 3 &= 0 \quad \text{or} \quad x - 3 = 0 \\
                x &= -3 \quad \text{or} \quad x = 3
            \end{align*}
        \end{example}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Practice: Advanced Factoring}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Solve Each of the Following]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item $x^2 + 7x + 12 = 0$
            \item $2x^2 - 5x - 3 = 0$
            \item $3x^2 + 10x + 7 = 0$
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Solutions: Advanced Factoring}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solutions]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item $x^2 + 7x + 12 = 0$
            \begin{align*}
                (x + 4)(x + 3) &= 0 \\
                x + 4 &= 0 \quad \text{or} \quad x + 3 = 0 \\
                x &= -4 \quad \text{or} \quad x = -3
            \end{align*}
            
            \item $2x^2 - 5x - 3 = 0$
            \begin{align*}
                (2x + 1)(x - 3) &= 0 \\
                2x + 1 &= 0 \quad \text{or} \quad x - 3 = 0 \\
                x &= -\frac{1}{2} \quad \text{or} \quad x = 3
            \end{align*}
            
            \item $3x^2 + 10x + 7 = 0$
            \begin{align*}
                (3x + 7)(x + 1) &= 0 \\
                3x + 7 &= 0 \quad \text{or} \quad x + 1 = 0 \\
                x &= -\frac{7}{3} \quad \text{or} \quad x = -1
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

% IV) WORD PROBLEMS
\begin{frame}{Word Problems}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Solve Each of the Following]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item The product of two consecutive numbers is 132. Find the numbers.
            
            \item The length of a rectangle is 5 meters more than twice its width. If the area is 75 square meters, find the dimensions.
            
            \item A 32m tall tree is broken during a storm. The distance from the base to the tip is 16m. At what height did the tree break?
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Word Problem 1}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Problem]
        \footnotesize
        The product of two consecutive numbers is 132. Find the numbers.
    \end{tcolorbox}
\end{frame}

\begin{frame}{Word Problem 1 - Solution}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution]
        \footnotesize
        The product of two consecutive numbers is 132. Find the numbers.
        
        \textbf{Solution:}
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Let $x$ be the first number
            \item Then $x + 1$ is the second number
            \item $x(x + 1) = 132$
            \item $x^2 + x - 132 = 0$
            \item $(x + 12)(x - 11) = 0$
            \item $x = -12$ or $x = 11$
        \end{enumerate}
        
        \textbf{Answer:} The numbers are 11 and 12, or -12 and -11
    \end{tcolorbox}
\end{frame}

\begin{frame}{Word Problem 2}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Problem]
        \footnotesize
        The length of a rectangle is 5 meters more than twice its width. If the area is 75 square meters, find the dimensions.
    \end{tcolorbox}
\end{frame}

\begin{frame}{Word Problem 2 - Solution}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution]
        \footnotesize
        The length of a rectangle is 5 meters more than twice its width. If the area is 75 square meters, find the dimensions.
        
        \textbf{Solution:}
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Let $x$ be the width
            \item Then $2x + 5$ is the length
            \item $x(2x + 5) = 75$
            \item $2x^2 + 5x - 75 = 0$
            \item $(2x + 15)(x - 5) = 0$
            \item $x = 5$ meters (width)
            \item Length = $2(5) + 5 = 15$ meters
        \end{enumerate}
        
        \textbf{Answer:} The dimensions are 5 meters by 15 meters
    \end{tcolorbox}
\end{frame}

\begin{frame}{Word Problem 3}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Problem]
        \footnotesize
        A 32m tall tree is broken during a storm. The distance from the base to the tip is 16m. At what height did the tree break?
    \end{tcolorbox}
\end{frame}

\begin{frame}{Word Problem 3 - Solution}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution]
        \footnotesize
        A 32m tall tree is broken during a storm. The distance from the base to the tip is 16m. At what height did the tree break?
        
        \textbf{Solution:}
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Let $x$ be the height where the tree broke
            \item Then $32 - x$ is the length of the fallen part
            \item Using Pythagorean theorem:
            \begin{align*}
                x^2 + 16^2 &= (32 - x)^2 \\
                x^2 + 256 &= 1024 - 64x + x^2 \\
                64x &= 768 \\
                x &= 12
            \end{align*}
        \end{enumerate}
        
        \textbf{Answer:} The tree broke at 12 meters
    \end{tcolorbox}
\end{frame}

% V) PRACTICE PROBLEMS
\begin{frame}{Practice Problems}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Solve Each of the Following]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item $2x^2 + 5x = 3$
            \item $3x^2 - 8x = -4$
            \item $4x^2 + 12x = -9$
            \item $x^2 - 6x = 16$
            \item $5x^2 + 7x = -2$
            \item $2x^2 - 9x = 5$
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Practice Problems - Solutions Part 1}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solutions]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item $2x^2 + 5x = 3$
            \begin{align*}
                2x^2 + 5x - 3 &= 0 \\
                (2x - 1)(x + 3) &= 0 \\
                x &= \frac{1}{2} \quad \text{or} \quad x = -3
            \end{align*}
            
            \item $3x^2 - 8x = -4$
            \begin{align*}
                3x^2 - 8x + 4 &= 0 \\
                (3x - 2)(x - 2) &= 0 \\
                x &= \frac{2}{3} \quad \text{or} \quad x = 2
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Practice Problems - Solutions Part 2}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solutions]
        \footnotesize
        \begin{enumerate}
            \setcounter{enumi}{2}
            \setlength{\itemsep}{0.5em}
            \item $4x^2 + 12x = -9$
            \begin{align*}
                4x^2 + 12x + 9 &= 0 \\
                (2x + 3)^2 &= 0 \\
                x &= -\frac{3}{2}
            \end{align*}
            
            \item $x^2 - 6x = 16$
            \begin{align*}
                x^2 - 6x - 16 &= 0 \\
                (x - 8)(x + 2) &= 0 \\
                x &= 8 \quad \text{or} \quad x = -2
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Practice Problems - Solutions Part 3}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solutions]
        \footnotesize
        \begin{enumerate}
            \setcounter{enumi}{4}
            \setlength{\itemsep}{0.5em}
            \item $5x^2 + 7x = -2$
            \begin{align*}
                5x^2 + 7x + 2 &= 0 \\
                (5x + 2)(x + 1) &= 0 \\
                x &= -\frac{2}{5} \quad \text{or} \quad x = -1
            \end{align*}
            
            \item $2x^2 - 9x = 5$
            \begin{align*}
                2x^2 - 9x - 5 &= 0 \\
                (2x + 1)(x - 5) &= 0 \\
                x &= -\frac{1}{2} \quad \text{or} \quad x = 5
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Summary}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Key Concepts]
        \footnotesize
        \begin{itemize}
            \item Understanding what it means to solve an equation
            \item Solving trinomials by factoring
            \item Steps for solving quadratic equations
            \item Word problems and applications
            \item Practice with various types of problems
        \end{itemize}
    \end{tcolorbox}
\end{frame}

% VI) CHALLENGE PROBLEMS
\begin{frame}{Challenge Problems}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Solve Each of the Following]
        \footnotesize
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item The difference of two numbers is 6. The sum of their squares is 90. Find the numbers.
            \item A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
            \item A rectangular field has an area of 120 square meters. Its length is 8 meters more than its width. Find the dimensions of the field.
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 1 - Solution Part 1}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 1)]
        \footnotesize
        The difference of two numbers is 6. The sum of their squares is 90. Find the numbers.
        
        \textbf{Solution:}
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Let the two numbers be $x$ and $y$.
            \item From the first condition: $x - y = 6 \Rightarrow y = x - 6$
            \item From the second condition: $x^2 + y^2 = 90$
            \item Substitute $y = x - 6$ into the second equation:
            \begin{align*}
                x^2 + (x - 6)^2 &= 90 \\
                x^2 + (x^2 - 12x + 36) &= 90 \\
                2x^2 - 12x + 36 - 90 &= 0 \\
                2x^2 - 12x - 54 &= 0
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 1 - Solution Part 2}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 2)]
        \footnotesize
        The difference of two numbers is 6. The sum of their squares is 90. Find the numbers.
        
        \textbf{Solution (Cont.):}
        \begin{enumerate}
            \setcounter{enumi}{4} % Continue numbering from previous part
            \setlength{\itemsep}{0.5em}
            \item Simplify and factor the quadratic equation:
            \begin{align*}
                x^2 - 6x - 27 &= 0 \\
                (x - 9)(x + 3) &= 0
            \end{align*}
            \item Solve for $x$:
            \begin{align*}
                x - 9 &= 0 \quad \text{or} \quad x + 3 = 0 \\
                x &= 9 \quad \text{or} \quad x = -3
            \end{align*}
            \item Find the corresponding values for $y$:
            \begin{itemize}
                \item If $x = 9$, then $y = 9 - 6 = 3$.
                \item If $x = -3$, then $y = -3 - 6 = -9$.
            \end{itemize}
        \end{enumerate}
        
        \textbf{Answer:} The numbers are (9, 3) or (-3, -9).
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 2}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Problem]
        \footnotesize
        A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 2 - Solution Part 1}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 1)]
        \footnotesize
        A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
        
        \textbf{Solution:}
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Let $w$ be the uniform width of the path.
            \item Original garden dimensions: Length = 10 m, Width = 7 m.
            \item Original garden area = $10 \times 7 = 70$ square meters.
            \item When a path of width $w$ is built around the garden, the new dimensions of the garden including the path will be:
            \begin{itemize}
                \item New Length = $10 + 2w$ (add $w$ to each side)
                \item New Width = $7 + 2w$ (add $w$ to each side)
            \end{itemize}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 2 - Solution Part 2}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 2)]
        \footnotesize
        A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
        
        \textbf{Solution (Cont.):}
        \begin{enumerate}
            \setcounter{enumi}{4} % Continue numbering from previous part
            \setlength{\itemsep}{0.5em}
            \item The total area (garden + path) is $(10 + 2w)(7 + 2w)$ square meters.
            \item The area of the path is the total area minus the original garden area.
            \item Area of path = $(10 + 2w)(7 + 2w) - 70$
            \item We are given that the area of the path is 54 square meters.
            \begin{align*}
                (10 + 2w)(7 + 2w) - 70 &= 54 \\
                70 + 20w + 14w + 4w^2 - 70 &= 54 \\
                4w^2 + 34w &= 54
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 2 - Solution Part 3}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 3)]
        \footnotesize
        A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
        
        \textbf{Solution (Cont.):}
        \begin{enumerate}
            \setcounter{enumi}{8} % Continue numbering from previous part
            \setlength{\itemsep}{0.5em}
            \item Rearrange the equation into standard quadratic form ($Aw^2 + Bw + C = 0$):
            \begin{align*}
                4w^2 + 34w - 54 &= 0
            \end{align*}
            \item Divide the entire equation by 2 to simplify (optional, but good practice):
            \begin{align*}
                2w^2 + 17w - 27 &= 0
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 2 - Solution Part 4}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 4)]
        \footnotesize
        A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
        
        \textbf{Solution (Cont.):}
        \begin{enumerate}
            \setcounter{enumi}{10} % Continue numbering from previous part
            \setlength{\itemsep}{0.5em}
            \item Use the quadratic formula to solve for $w$:
            \begin{align*}
                w &= \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} \\
                w &= \frac{-17 \pm \sqrt{17^2 - 4(2)(-27)}}{2(2)} \\
                w &= \frac{-17 \pm \sqrt{289 + 216}}{4} \\
                w &= \frac{-17 \pm \sqrt{505}}{4}
            \end{align*}
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 2 - Solution Part 5}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 5)]
        \footnotesize
        A rectangular garden is 10 meters long and 7 meters wide. A path of uniform width is to be built around the garden. If the area of the path is 54 square meters, find the width of the path.
        
        \textbf{Solution (Cont.):}
        \begin{enumerate}
            \setcounter{enumi}{11} % Continue numbering from previous part
            \setlength{\itemsep}{0.5em}
            \item Calculate the two possible values for $w$:
            \begin{align*}
                w_1 &= \frac{-17 + \sqrt{505}}{4} \approx \frac{-17 + 22.47}{4} \approx \frac{5.47}{4} \approx 1.3675 \\
                w_2 &= \frac{-17 - \sqrt{505}}{4} \approx \frac{-17 - 22.47}{4} \approx \frac{-39.47}{4} \approx -9.8675
            \end{align*}
            \item Since the width cannot be negative, we discard $w_2 \approx -9.8675$.
        \end{enumerate}
        
        \textbf{Answer:} The width of the path is approximately 1.37 meters.
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 3}
    \begin{tcolorbox}[colback=lightgray,colframe=primary,title=Problem]
        \footnotesize
        A rectangular field has an area of 120 square meters. Its length is 8 meters more than its width. Find the dimensions of the field.
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 3 - Solution Part 1}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 1)]
        \footnotesize
        A rectangular field has an area of 120 square meters. Its length is 8 meters more than its width. Find the dimensions of the field.
        
        \textbf{Solution:}
        \begin{enumerate}
            \setlength{\itemsep}{0.5em}
            \item Let the width of the field be $x$ meters.
            \item The length of the field is $x + 8$ meters.
            \item The area of the field is given by Length $\times$ Width, so $x(x + 8) = 120$.
            \item Expand and rearrange the equation into standard quadratic form:
            \begin{align*}
                x^2 + 8x &= 120 \\
                x^2 + 8x - 120 &= 0
            \end{align*}
            \item Identify the coefficients for the quadratic formula: $a=1$, $b=8$, $c=-120$.
        \end{enumerate}
    \end{tcolorbox}
\end{frame}

\begin{frame}{Challenge Problem 3 - Solution Part 2}
    \begin{tcolorbox}[colback=lightgray,colframe=accent,title=Detailed Solution (Part 2)]
        \footnotesize
        A rectangular field has an area of 120 square meters. Its length is 8 meters more than its width. Find the dimensions of the field.
        
        \textbf{Solution (Cont.):}
        \begin{enumerate}
            \setcounter{enumi}{5} % Continue numbering from previous part
            \setlength{\itemsep}{0.5em}
            \item Use the quadratic formula to solve for $x$:
            \begin{align*}
                x &= \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} \\
                x &= \frac{-8 \pm \sqrt{8^2 - 4(1)(-120)}}{2(1)} \\
                x &= \frac{-8 \pm \sqrt{64 + 480}}{2} \\
                x &= \frac{-8 \pm \sqrt{544}}{2}
            \end{align*}
            \item Simplify the square root: $\sqrt{544} = \sqrt{16 \times 34} = 4\sqrt{34}$.
            \begin{align*}
                x &= \frac{-8 \pm 4\sqrt{34}}{2} \\
                x &= -4 \pm 2\sqrt{34}
            \end{align*}
            \item Calculate the two possible values for $x$ (approximately):
            \begin{itemize}
                \item $x_1 = -4 + 2\sqrt{34} \approx -4 + 2(5.83) \approx -4 + 11.66 \approx 7.66$
                \item $x_2 = -4 - 2\sqrt{34} \approx -4 - 2(5.83) \approx -4 - 11.66 \approx -15.66$
            \end{itemize}
            \item Since the width cannot be negative, we discard $x_2$.
        \end{enumerate}
        
        \textbf{Answer:} The width of the field is approximately 7.66 meters, and the length is approximately $7.66 + 8 = 15.66$ meters.
    \end{tcolorbox}
\end{frame}

\end{document} 