% !TEX program = pdflatex
\documentclass[aspectratio=169]{beamer}
\usetheme{Madrid}
\usecolortheme{default}
\usepackage{tikz}

% Custom colors
\definecolor{myblue}{RGB}{0,114,178}
\definecolor{myorange}{RGB}{230,159,0}
\definecolor{mygreen}{RGB}{0,158,115}
\definecolor{myred}{RGB}{213,94,0}
\definecolor{mypurple}{RGB}{204,121,167}
\definecolor{mybrown}{RGB}{240,228,66}
\definecolor{mygray}{RGB}{189,189,189}

% Set colors
\setbeamercolor{title}{fg=white,bg=myblue}
\setbeamercolor{frametitle}{fg=white,bg=myblue}
\setbeamercolor{section title}{fg=white,bg=myblue}
\setbeamercolor{subsection title}{fg=white,bg=myorange}
\setbeamercolor{item}{fg=myblue}
\setbeamercolor{subitem}{fg=myorange}

% Remove navigation symbols
\setbeamertemplate{navigation symbols}{}

% Custom commands
\newcommand{\limx}[2]{\lim_{x \to #1} #2}

% Title page
\title{Related Rates}
\subtitle{Applications of Derivatives in Real-World Problems}
\author{Differential Calculus}
\date{}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}{What are Related Rates?}
\begin{itemize}
  \item Problems where you are given the rate of change of one quantity
  \item You need to determine the rate of change of another related quantity
  \item All quantities are functions of time
  \item Use implicit differentiation to relate the rates
\end{itemize}
\end{frame}

\begin{frame}{General Strategy}
\begin{block}{Step-by-Step Method}
\begin{enumerate}
  \item \textbf{Draw a picture} and identify all relevant variables
  \item \textbf{Write an equation} relating the quantities
  \item \textbf{Differentiate both sides} with respect to time
  \item \textbf{Substitute known values} and solve for the unknown rate
\end{enumerate}
\end{block}
\end{frame}

\begin{frame}{Key Concepts}
\begin{block}{Important Points}
\begin{itemize}
  \item All variables are functions of time: $x(t)$, $y(t)$, $V(t)$, etc.
  \item Use implicit differentiation: $\frac{d}{dt}[f(x,y)] = \frac{\partial f}{\partial x}\frac{dx}{dt} + \frac{\partial f}{\partial y}\frac{dy}{dt}$
  \item Pay attention to units and signs
  \item Draw diagrams to visualize the problem
\end{itemize}
\end{block}
\end{frame}

\section{Basic Examples}

\begin{frame}{Example 1: Spherical Balloon}
\textbf{Problem:} A spherical balloon is being inflated at a rate of $13$ cm$^3$/sec. How fast is the radius changing when the balloon has radius $15$ cm?
\end{frame}

\begin{frame}{Solution: Spherical Balloon}
\textbf{Solution:}
\[
\begin{aligned}
  V &= \frac{4}{3}\pi r^3 \\
  \frac{dV}{dt} &= 4\pi r^2 \cdot \frac{dr}{dt} \\
  \frac{dr}{dt} &= \frac{1}{4\pi r^2} \cdot \frac{dV}{dt} \\
  \frac{dr}{dt} &= \frac{13}{4\pi (15)^2} \\
  \frac{dr}{dt} &= \frac{13}{900\pi} \text{ cm/sec}
\end{aligned}
\]
\end{frame}

\begin{frame}{Example 2: Ladder Problem}
\textbf{Problem:} A 5m ladder is leaning against a wall. The base slides out at 1m/s. How fast is the top sliding down when the base is 3m from the wall?
\end{frame}

\begin{frame}{Solution: Ladder Problem}
\textbf{Solution:}
\[
\begin{aligned}
  x^2 + y^2 &= 25 \\
  2x\frac{dx}{dt} + 2y\frac{dy}{dt} &= 0 \\
  x\frac{dx}{dt} + y\frac{dy}{dt} &= 0 \\
  \frac{dy}{dt} &= -\frac{x}{y}\frac{dx}{dt}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution: Ladder Problem (continued)}
\textbf{Solution (continued):}
\[
\begin{aligned}
  \text{When } x &= 3, y = \sqrt{25 - 9} = 4 \\
  \frac{dy}{dt} &= -\frac{3}{4} \cdot 1 \\
  \frac{dy}{dt} &= -\frac{3}{4} \text{ m/s}
\end{aligned}
\]
\end{frame}

\section{Intermediate Examples}

\begin{frame}{Example 3: Balloon Angle}
\textbf{Problem:} A balloon rises vertically from a point 200m away. When the angle is $\pi/4$, it's changing at 0.05 rad/sec. How fast is the balloon rising?
\end{frame}

\begin{frame}{Solution: Balloon Angle}
\textbf{Solution:}
\[
\begin{aligned}
  h &= 200 \tan \theta \\
  \frac{dh}{dt} &= 200 \sec^2 \theta \cdot \frac{d\theta}{dt} \\
  \frac{dh}{dt} &= 200 \cdot (\sqrt{2})^2 \cdot 0.05 \\
  \frac{dh}{dt} &= 200 \cdot 2 \cdot 0.05 \\
  \frac{dh}{dt} &= 20 \text{ m/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Example 4: Percentage Rates}
\textbf{Problem:} $R = PQ$. $P$ increases at 8\%/year, $Q$ decreases at 2\%/year. What's the percentage rate of change for $R$?
\end{frame}

\begin{frame}{Solution: Percentage Rates}
\textbf{Solution:}
\[
\begin{aligned}
  R &= PQ \\
  \frac{dR}{dt} &= P\frac{dQ}{dt} + Q\frac{dP}{dt} \\
  \frac{100}{R}\frac{dR}{dt} &= \frac{100P}{R}\frac{dQ}{dt} + \frac{100Q}{R}\frac{dP}{dt} \\
  &= \frac{100}{Q}\frac{dQ}{dt} + \frac{100}{P}\frac{dP}{dt} \\
  &= -2 + 8 = 6\%
\end{aligned}
\]
\end{frame}

\section{Advanced Examples}

\begin{frame}{Example 5: Shadow Problem}
\textbf{Problem:} A ball drops from 49m height. A light 49m high is 10m to the left. How fast is the shadow moving 1 second after drop?
\end{frame}

\begin{frame}{Solution: Shadow Problem}
\textbf{Solution:}
\[
\begin{aligned}
  h(t) &= 49 - 4.9t^2 \\
  \text{By similar triangles: } \frac{4.9t^2}{10} &= \frac{49 - 4.9t^2}{s(t)} \\
  s(t) &= \frac{10(49 - 4.9t^2)}{4.9t^2} = \frac{100}{t^2} - 10 \\
  \frac{ds}{dt} &= -\frac{200}{t^3} \\
  \frac{ds}{dt}(1) &= -200 \text{ m/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Example 6: Boats Problem}
\textbf{Problem:} Boat A is 15km west of boat B. A travels east at 3km/h, B travels north at 4km/h. How fast is distance changing at 3pm?
\end{frame}

\begin{frame}{Solution: Boats Problem}
\textbf{Solution:}
\[
\begin{aligned}
  z^2 &= x^2 + y^2 \\
  2z\frac{dz}{dt} &= 2x\frac{dx}{dt} + 2y\frac{dy}{dt} \\
  \frac{dz}{dt} &= \frac{x\frac{dx}{dt} + y\frac{dy}{dt}}{z}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution: Boats Problem (continued)}
\textbf{Solution (continued):}
\[
\begin{aligned}
  \text{At 3pm: } x &= 6, y = 12, z = \sqrt{180} = 6\sqrt{5} \\
  \frac{dx}{dt} &= -3, \frac{dy}{dt} = 4 \\
  \frac{dz}{dt} &= \frac{6(-3) + 12(4)}{6\sqrt{5}} \\
  \frac{dz}{dt} &= \frac{30}{6\sqrt{5}} = \sqrt{5} \text{ km/h}
\end{aligned}
\]
\end{frame}

\section{Basic Practice Problems}

\begin{frame}{Practice: 1 and 2}
\textbf{Practice 1:} \\
A spherical balloon is inflated at 8 cm$^3$/sec. \\
How fast is radius changing when radius is 10 cm? \\
\vspace{1em}
\textbf{Practice 2:} \\
A 13m ladder slides down a wall. Base moves at 2m/s. \\
How fast is top sliding when base is 5m from wall?
\end{frame}

\begin{frame}{Practice: 3 and 4}
\textbf{Practice 3:} \\
A cone's height increases at 3cm/s. \\
How fast is volume changing when height is 10cm and radius is 5cm? \\
\vspace{1em}
\textbf{Practice 4:} \\
A rectangle's length increases at 2cm/s, width decreases at 1cm/s. \\
How fast is area changing when length is 8cm and width is 6cm?
\end{frame}

\begin{frame}{Practice: 5 and 6}
\textbf{Practice 5:} \\
A cube's edge length increases at 4cm/s. \\
How fast is volume changing when edge length is 12cm? \\
\vspace{1em}
\textbf{Practice 6:} \\
A circle's radius increases at 2cm/s. \\
How fast is area changing when radius is 7cm?
\end{frame}

\begin{frame}{Practice: 7 and 8}
\textbf{Practice 7:} \\
A cylinder's radius increases at 1cm/s, height decreases at 2cm/s. \\
How fast is volume changing when radius is 6cm and height is 10cm? \\
\vspace{1em}
\textbf{Practice 8:} \\
A square's side length increases at 3cm/s. \\
How fast is diagonal changing when side length is 9cm?
\end{frame}

\section{Intermediate Practice Problems}

\begin{frame}{Practice: 9 and 10}
\textbf{Practice 9:} \\
A person walks away from a 20m light pole at 2m/s. \\
How fast is shadow length changing when person is 15m from pole? \\
\vspace{1em}
\textbf{Practice 10:} \\
A plane flies at 500km/h at altitude 10km. \\
How fast is distance from observer changing when plane is 15km away horizontally?
\end{frame}

\begin{frame}{Practice: 11 and 12}
\textbf{Practice 11:} \\
A conical tank has radius 5m, height 10m. \\
Water flows in at 3m$^3$/min. \\
How fast is water level rising when depth is 4m? \\
\vspace{1em}
\textbf{Practice 12:} \\
A street light is 8m high. A person 1.8m tall walks away at 1.5m/s. \\
How fast is shadow length changing when person is 6m from light?
\end{frame}

\begin{frame}{Practice: 13 and 14}
\textbf{Practice 13:} \\
A kite flies at height 100m. String is let out at 2m/s. \\
How fast is kite moving horizontally when 200m of string is out? \\
\vspace{1em}
\textbf{Practice 14:} \\
A balloon rises at 3m/s. Observer 50m away. \\
How fast is angle of elevation changing when balloon is 30m high?
\end{frame}

\begin{frame}{Practice: 15 and 16}
\textbf{Practice 15:} \\
A rectangular pool is 20m long, 10m wide. \\
Water flows in at 5m$^3$/min. \\
How fast is water level rising when depth is 2m? \\
\vspace{1em}
\textbf{Practice 16:} \\
A particle moves along curve $y = x^2$ at 2 units/sec. \\
How fast is distance from origin changing when $x = 3$?
\end{frame}

\section{Advanced Practice Problems}

\begin{frame}{Practice: 17 and 18}
\textbf{Practice 17:} \\
A cylindrical tank radius 4m, height 8m, lying on side. \\
Fuel flows in at 2m$^3$/min. \\
How fast is fuel level rising when depth is 3m? \\
\vspace{1em}
\textbf{Practice 18:} \\
Two cars approach intersection. Car A at 60km/h, Car B at 80km/h. \\
How fast is distance changing when A is 3km away, B is 4km away?
\end{frame}

\begin{frame}{Practice: 19 and 20}
\textbf{Practice 19:} \\
A conical pile of sand has height equal to radius. \\
Sand falls at 10m$^3$/min. \\
How fast is radius changing when height is 5m? \\
\vspace{1em}
\textbf{Practice 20:} \\
A ladder 10m long slides down wall. Base moves at 1.5m/s. \\
How fast is angle between ladder and ground changing when base is 6m from wall?
\end{frame}

\begin{frame}{Practice: 21 and 22}
\textbf{Practice 21:} \\
A spherical snowball melts at rate proportional to surface area. \\
How fast is radius changing when radius is 8cm? \\
\vspace{1em}
\textbf{Practice 22:} \\
A particle moves along $y = \sin x$ at 3 units/sec. \\
How fast is distance from $(0,0)$ changing when $x = \pi/4$?
\end{frame}

\begin{frame}{Practice: 23 and 24}
\textbf{Practice 23:} \\
A rectangular box has length, width, height increasing at 2cm/s, 1cm/s, 3cm/s. \\
How fast is volume changing when dimensions are 8cm, 6cm, 4cm? \\
\vspace{1em}
\textbf{Practice 24:} \\
A circular oil slick expands at 2m$^2$/min. \\
How fast is radius changing when area is $100\pi$ m$^2$?
\end{frame}

\begin{frame}{Practice: 25}
\textbf{Practice 25:} \\
A conical cup has radius 5cm, height 12cm. \\
Water flows out at 3cm$^3$/s. \\
How fast is water level falling when depth is 6cm?
\end{frame}

\section{Solutions to Basic Practice Problems}

\begin{frame}{Solution to Practice 1}
\textbf{Practice 1:}

A spherical balloon is inflated at 8 cm$^3$/sec. How fast is radius changing when radius is 10 cm?

\textbf{Solution:}
\[
\begin{aligned}
  V &= \frac{4}{3}\pi r^3 \\
  \frac{dV}{dt} &= 4\pi r^2 \cdot \frac{dr}{dt} \\
  \frac{dr}{dt} &= \frac{1}{4\pi r^2} \cdot \frac{dV}{dt} \\
  \frac{dr}{dt} &= \frac{8}{4\pi (10)^2} \\
  \frac{dr}{dt} &= \frac{8}{400\pi} = \frac{1}{50\pi} \text{ cm/sec}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 2}
\textbf{Practice 2:}

A 13m ladder slides down a wall. Base moves at 2m/s. How fast is top sliding when base is 5m from wall?

\textbf{Solution:}
\[
\begin{aligned}
  x^2 + y^2 &= 169 \\
  2x\frac{dx}{dt} + 2y\frac{dy}{dt} &= 0 \\
  x\frac{dx}{dt} + y\frac{dy}{dt} &= 0 \\
  \frac{dy}{dt} &= -\frac{x}{y}\frac{dx}{dt}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 2 (continued)}
\textbf{Solution (continued):}
\[
\begin{aligned}
  \text{When } x &= 5, y = \sqrt{169 - 25} = 12 \\
  \frac{dy}{dt} &= -\frac{5}{12} \cdot 2 \\
  \frac{dy}{dt} &= -\frac{10}{12} = -\frac{5}{6} \text{ m/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 3}
\textbf{Practice 3:}

A cone's height increases at 3cm/s. How fast is volume changing when height is 10cm and radius is 5cm?

\textbf{Solution:}
\[
\begin{aligned}
  V &= \frac{1}{3}\pi r^2 h \\
  \frac{dV}{dt} &= \frac{1}{3}\pi(2r\frac{dr}{dt} \cdot h + r^2\frac{dh}{dt}) \\
  \text{Since } \frac{dr}{dt} &= 0 \text{ (radius constant)} \\
  \frac{dV}{dt} &= \frac{1}{3}\pi r^2 \frac{dh}{dt} \\
  \frac{dV}{dt} &= \frac{1}{3}\pi(5)^2 \cdot 3 \\
  \frac{dV}{dt} &= 25\pi \text{ cm}^3\text{/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 4}
\textbf{Practice 4:}

A rectangle's length increases at 2cm/s, width decreases at 1cm/s. How fast is area changing when length is 8cm and width is 6cm?

\textbf{Solution:}
\[
\begin{aligned}
  A &= lw \\
  \frac{dA}{dt} &= l\frac{dw}{dt} + w\frac{dl}{dt} \\
  \frac{dA}{dt} &= 8(-1) + 6(2) \\
  \frac{dA}{dt} &= -8 + 12 = 4 \text{ cm}^2\text{/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 5}
\textbf{Practice 5:}

A cube's edge length increases at 4cm/s. How fast is volume changing when edge length is 12cm?

\textbf{Solution:}
\[
\begin{aligned}
  V &= s^3 \\
  \frac{dV}{dt} &= 3s^2 \cdot \frac{ds}{dt} \\
  \frac{dV}{dt} &= 3(12)^2 \cdot 4 \\
  \frac{dV}{dt} &= 3 \cdot 144 \cdot 4 = 1728 \text{ cm}^3\text{/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 6}
\textbf{Practice 6:}

A circle's radius increases at 2cm/s. How fast is area changing when radius is 7cm?

\textbf{Solution:}
\[
\begin{aligned}
  A &= \pi r^2 \\
  \frac{dA}{dt} &= 2\pi r \cdot \frac{dr}{dt} \\
  \frac{dA}{dt} &= 2\pi(7) \cdot 2 \\
  \frac{dA}{dt} &= 28\pi \text{ cm}^2\text{/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 7}
\textbf{Practice 7:}

A cylinder's radius increases at 1cm/s, height decreases at 2cm/s. How fast is volume changing when radius is 6cm and height is 10cm?

\textbf{Solution:}
\[
\begin{aligned}
  V &= \pi r^2 h \\
  \frac{dV}{dt} &= \pi(2r\frac{dr}{dt} \cdot h + r^2\frac{dh}{dt}) \\
  \frac{dV}{dt} &= \pi(2 \cdot 6 \cdot 1 \cdot 10 + 6^2 \cdot (-2)) \\
  \frac{dV}{dt} &= \pi(120 - 72) = 48\pi \text{ cm}^3\text{/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 8}
\textbf{Practice 8:}

A square's side length increases at 3cm/s. How fast is diagonal changing when side length is 9cm?

\textbf{Solution:}
\[
\begin{aligned}
  d &= s\sqrt{2} \\
  \frac{dd}{dt} &= \sqrt{2} \cdot \frac{ds}{dt} \\
  \frac{dd}{dt} &= \sqrt{2} \cdot 3 = 3\sqrt{2} \text{ cm/s}
\end{aligned}
\]
\end{frame}

\section{Solutions to Intermediate Practice Problems}

\begin{frame}{Solution to Practice 9}
\textbf{Practice 9:}

A person walks away from a 20m light pole at 2m/s. How fast is shadow length changing when person is 15m from pole?

\textbf{Solution:}
\[
\begin{aligned}
  \text{By similar triangles: } \frac{h}{x} &= \frac{20}{x + s} \\
  h(x + s) &= 20x \\
  hs &= 20x - hx \\
  s &= \frac{20x - hx}{h} = \frac{(20 - h)x}{h}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 9 (continued)}
\textbf{Solution (continued):}
\[
\begin{aligned}
  \frac{ds}{dt} &= \frac{(20 - h)}{h} \cdot \frac{dx}{dt} \\
  \text{Assuming person height } h &= 1.8m \\
  \frac{ds}{dt} &= \frac{18.2}{1.8} \cdot 2 \\
  \frac{ds}{dt} &= \frac{36.4}{1.8} \approx 20.2 \text{ m/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 10}
\textbf{Practice 10:}

A plane flies at 500km/h at altitude 10km. How fast is distance from observer changing when plane is 15km away horizontally?

\textbf{Solution:}
\[
\begin{aligned}
  d^2 &= x^2 + h^2 \\
  2d\frac{dd}{dt} &= 2x\frac{dx}{dt} \\
  \frac{dd}{dt} &= \frac{x}{d}\frac{dx}{dt} \\
  d &= \sqrt{15^2 + 10^2} = \sqrt{325} \\
  \frac{dd}{dt} &= \frac{15}{\sqrt{325}} \cdot 500 \\
  \frac{dd}{dt} &= \frac{7500}{\sqrt{325}} \text{ km/h}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 11}
\textbf{Practice 11:}

A conical tank has radius 5m, height 10m. Water flows in at 3m$^3$/min. How fast is water level rising when depth is 4m?

\textbf{Solution:}
\[
\begin{aligned}
  \frac{r}{h} &= \frac{5}{10} = \frac{1}{2} \\
  r &= \frac{h}{2} \\
  V &= \frac{1}{3}\pi r^2 h = \frac{1}{3}\pi(\frac{h}{2})^2 h = \frac{\pi h^3}{12} \\
  \frac{dV}{dt} &= \frac{\pi h^2}{4} \cdot \frac{dh}{dt} \\
  \frac{dh}{dt} &= \frac{4}{\pi h^2} \cdot \frac{dV}{dt} \\
  \frac{dh}{dt} &= \frac{4}{\pi(4)^2} \cdot 3 = \frac{12}{16\pi} = \frac{3}{4\pi} \text{ m/min}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 12}
\textbf{Practice 12:}

A street light is 8m high. A person 1.8m tall walks away at 1.5m/s. How fast is shadow length changing when person is 6m from light?

\textbf{Solution:}
\[
\begin{aligned}
  \text{By similar triangles: } \frac{1.8}{s} &= \frac{8}{x + s} \\
  1.8(x + s) &= 8s \\
  1.8x + 1.8s &= 8s \\
  1.8x &= 6.2s \\
  s &= \frac{1.8x}{6.2} = \frac{9x}{31}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 12 (continued)}
\textbf{Solution (continued):}
\[
\begin{aligned}
  \frac{ds}{dt} &= \frac{9}{31} \cdot \frac{dx}{dt} \\
  \frac{ds}{dt} &= \frac{9}{31} \cdot 1.5 \\
  \frac{ds}{dt} &= \frac{13.5}{31} \approx 0.44 \text{ m/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 13}
\textbf{Practice 13:}

A kite flies at height 100m. String is let out at 2m/s. How fast is kite moving horizontally when 200m of string is out?

\textbf{Solution:}
\[
\begin{aligned}
  x^2 + 100^2 &= 200^2 \\
  x^2 &= 40000 - 10000 = 30000 \\
  x &= \sqrt{30000} = 100\sqrt{3} \\
  2x\frac{dx}{dt} &= 2s\frac{ds}{dt} \\
  \frac{dx}{dt} &= \frac{s}{x}\frac{ds}{dt} \\
  \frac{dx}{dt} &= \frac{200}{100\sqrt{3}} \cdot 2 = \frac{400}{100\sqrt{3}} = \frac{4}{\sqrt{3}} \text{ m/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 14}
\textbf{Practice 14:}

A balloon rises at 3m/s. Observer 50m away. How fast is angle of elevation changing when balloon is 30m high?

\textbf{Solution:}
\[
\begin{aligned}
  \tan \theta &= \frac{h}{50} \\
  \sec^2 \theta \cdot \frac{d\theta}{dt} &= \frac{1}{50} \cdot \frac{dh}{dt} \\
  \frac{d\theta}{dt} &= \frac{1}{50 \sec^2 \theta} \cdot 3 \\
  \sec \theta &= \frac{\sqrt{50^2 + 30^2}}{50} = \frac{\sqrt{3400}}{50} = \frac{\sqrt{34}}{5} \\
  \frac{d\theta}{dt} &= \frac{3}{50 \cdot \frac{34}{25}} = \frac{75}{1700} = \frac{3}{68} \text{ rad/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 15}
\textbf{Practice 15:}

A rectangular pool is 20m long, 10m wide. Water flows in at 5m$^3$/min. How fast is water level rising when depth is 2m?

\textbf{Solution:}
\[
\begin{aligned}
  V &= lwh \\
  \frac{dV}{dt} &= lw\frac{dh}{dt} \\
  \frac{dh}{dt} &= \frac{1}{lw} \cdot \frac{dV}{dt} \\
  \frac{dh}{dt} &= \frac{1}{20 \cdot 10} \cdot 5 \\
  \frac{dh}{dt} &= \frac{5}{200} = \frac{1}{40} \text{ m/min}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 16}
\textbf{Practice 16:}

A particle moves along curve $y = x^2$ at 2 units/sec. How fast is distance from origin changing when $x = 3$?

\textbf{Solution:}
\[
\begin{aligned}
  d^2 &= x^2 + y^2 = x^2 + x^4 \\
  2d\frac{dd}{dt} &= 2x\frac{dx}{dt} + 4x^3\frac{dx}{dt} \\
  \frac{dd}{dt} &= \frac{x + 2x^3}{d}\frac{dx}{dt} \\
  \text{When } x &= 3, y = 9, d = \sqrt{9 + 81} = \sqrt{90} = 3\sqrt{10} \\
  \frac{dd}{dt} &= \frac{3 + 2(27)}{3\sqrt{10}} \cdot 2 = \frac{57}{3\sqrt{10}} \cdot 2 = \frac{38}{\sqrt{10}} \text{ units/sec}
\end{aligned}
\]
\end{frame}

\section{Solutions to Advanced Practice Problems}

\begin{frame}{Solution to Practice 17}
\textbf{Practice 17:}

A cylindrical tank radius 4m, height 8m, lying on side. Fuel flows in at 2m$^3$/min. How fast is fuel level rising when depth is 3m?

\textbf{Solution:}
\[
\begin{aligned}
  V &= L \cdot \text{Area of segment} \\
  \text{Area} &= r^2\theta - \frac{r^2}{2}\sin(2\theta) \\
  h &= r - r\cos\theta \\
  \cos\theta &= 1 - \frac{h}{r} = 1 - \frac{3}{4} = \frac{1}{4} \\
  \theta &= \arccos(\frac{1}{4})
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 17 (continued)}
\textbf{Solution (continued):}
\[
\begin{aligned}
  \frac{dV}{dt} &= Lr^2(1 - \cos(2\theta))\frac{d\theta}{dt} \\
  \frac{dh}{dt} &= r\sin\theta \cdot \frac{d\theta}{dt} \\
  \frac{dh}{dt} &= r\sin\theta \cdot \frac{2}{Lr^2(1 - \cos(2\theta))} \\
  \frac{dh}{dt} &= \frac{2\sin\theta}{Lr(1 - \cos(2\theta))} \\
  \frac{dh}{dt} &= \frac{2 \cdot \frac{\sqrt{15}}{4}}{8 \cdot 4 \cdot 2 \cdot (\frac{\sqrt{15}}{4})^2} = \frac{1}{16\sqrt{15}} \text{ m/min}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 18}
\textbf{Practice 18:}

Two cars approach intersection. Car A at 60km/h, Car B at 80km/h. How fast is distance changing when A is 3km away, B is 4km away?

\textbf{Solution:}
\[
\begin{aligned}
  d^2 &= x^2 + y^2 \\
  2d\frac{dd}{dt} &= 2x\frac{dx}{dt} + 2y\frac{dy}{dt} \\
  \frac{dd}{dt} &= \frac{x\frac{dx}{dt} + y\frac{dy}{dt}}{d} \\
  d &= \sqrt{3^2 + 4^2} = 5 \\
  \frac{dd}{dt} &= \frac{3(-60) + 4(-80)}{5} \\
  \frac{dd}{dt} &= \frac{-180 - 320}{5} = -100 \text{ km/h}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 19}
\textbf{Practice 19:}

A conical pile of sand has height equal to radius. Sand falls at 10m$^3$/min. How fast is radius changing when height is 5m?

\textbf{Solution:}
\[
\begin{aligned}
  h &= r \\
  V &= \frac{1}{3}\pi r^2 h = \frac{1}{3}\pi r^3 \\
  \frac{dV}{dt} &= \pi r^2 \cdot \frac{dr}{dt} \\
  \frac{dr}{dt} &= \frac{1}{\pi r^2} \cdot \frac{dV}{dt} \\
  \frac{dr}{dt} &= \frac{10}{\pi(5)^2} = \frac{10}{25\pi} = \frac{2}{5\pi} \text{ m/min}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 20}
\textbf{Practice 20:}

A ladder 10m long slides down wall. Base moves at 1.5m/s. How fast is angle between ladder and ground changing when base is 6m from wall?

\textbf{Solution:}
\[
\begin{aligned}
  \cos\theta &= \frac{x}{10} \\
  -\sin\theta \cdot \frac{d\theta}{dt} &= \frac{1}{10} \cdot \frac{dx}{dt} \\
  \frac{d\theta}{dt} &= -\frac{1}{10\sin\theta} \cdot \frac{dx}{dt} \\
  \text{When } x &= 6, \sin\theta = \frac{8}{10} = \frac{4}{5} \\
  \frac{d\theta}{dt} &= -\frac{1}{10 \cdot \frac{4}{5}} \cdot 1.5 = -\frac{5}{40} \cdot 1.5 = -\frac{3}{16} \text{ rad/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 21}
\textbf{Practice 21:}

A spherical snowball melts at rate proportional to surface area. How fast is radius changing when radius is 8cm?

\textbf{Solution:}
\[
\begin{aligned}
  \frac{dV}{dt} &= -k \cdot 4\pi r^2 \\
  V &= \frac{4}{3}\pi r^3 \\
  \frac{dV}{dt} &= 4\pi r^2 \cdot \frac{dr}{dt} \\
  4\pi r^2 \cdot \frac{dr}{dt} &= -k \cdot 4\pi r^2 \\
  \frac{dr}{dt} &= -k \text{ (constant rate)}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 22}
\textbf{Practice 22:}

A particle moves along $y = \sin x$ at 3 units/sec. How fast is distance from $(0,0)$ changing when $x = \pi/4$?

\textbf{Solution:}
\[
\begin{aligned}
  d^2 &= x^2 + y^2 = x^2 + \sin^2 x \\
  2d\frac{dd}{dt} &= 2x\frac{dx}{dt} + 2\sin x \cos x \cdot \frac{dx}{dt} \\
  \frac{dd}{dt} &= \frac{x + \sin x \cos x}{d} \cdot \frac{dx}{dt} \\
  \text{When } x &= \pi/4, y = \sin(\pi/4) = \frac{\sqrt{2}}{2} \\
  d &= \sqrt{(\pi/4)^2 + (\sqrt{2}/2)^2} = \sqrt{\pi^2/16 + 1/2} \\
  \frac{dd}{dt} &= \frac{\pi/4 + \frac{\sqrt{2}}{2} \cdot \frac{\sqrt{2}}{2}}{d} \cdot 3 = \frac{\pi/4 + 1/2}{d} \cdot 3
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 23}
\textbf{Practice 23:}

A rectangular box has length, width, height increasing at 2cm/s, 1cm/s, 3cm/s. How fast is volume changing when dimensions are 8cm, 6cm, 4cm?

\textbf{Solution:}
\[
\begin{aligned}
  V &= lwh \\
  \frac{dV}{dt} &= lw\frac{dh}{dt} + lh\frac{dw}{dt} + wh\frac{dl}{dt} \\
  \frac{dV}{dt} &= 8 \cdot 6 \cdot 3 + 8 \cdot 4 \cdot 1 + 6 \cdot 4 \cdot 2 \\
  \frac{dV}{dt} &= 144 + 32 + 48 = 224 \text{ cm}^3\text{/s}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 24}
\textbf{Practice 24:}

A circular oil slick expands at 2m$^2$/min. How fast is radius changing when area is $100\pi$ m$^2$?

\textbf{Solution:}
\[
\begin{aligned}
  A &= \pi r^2 \\
  \frac{dA}{dt} &= 2\pi r \cdot \frac{dr}{dt} \\
  \frac{dr}{dt} &= \frac{1}{2\pi r} \cdot \frac{dA}{dt} \\
  \text{When } A &= 100\pi, r = 10 \\
  \frac{dr}{dt} &= \frac{2}{2\pi \cdot 10} = \frac{1}{10\pi} \text{ m/min}
\end{aligned}
\]
\end{frame}

\begin{frame}{Solution to Practice 25}
\textbf{Practice 25:}

A conical cup has radius 5cm, height 12cm. Water flows out at 3cm$^3$/s. How fast is water level falling when depth is 6cm?

\textbf{Solution:}
\[
\begin{aligned}
  \frac{r}{h} &= \frac{5}{12} \\
  r &= \frac{5h}{12} \\
  V &= \frac{1}{3}\pi r^2 h = \frac{1}{3}\pi(\frac{5h}{12})^2 h = \frac{25\pi h^3}{432} \\
  \frac{dV}{dt} &= \frac{25\pi h^2}{144} \cdot \frac{dh}{dt} \\
  \frac{dh}{dt} &= \frac{144}{25\pi h^2} \cdot \frac{dV}{dt} \\
  \frac{dh}{dt} &= \frac{144}{25\pi(6)^2} \cdot (-3) = -\frac{432}{900\pi} = -\frac{12}{25\pi} \text{ cm/s}
\end{aligned}
\]
\end{frame}

\section{Summary}

\begin{frame}{Key Points - Related Rates}
\begin{itemize}
    \item \textbf{Strategy:} Draw picture, write equation, differentiate, substitute
    \item \textbf{Key concept:} All variables are functions of time
    \item \textbf{Method:} Use implicit differentiation with respect to time
    \item \textbf{Units:} Pay attention to units and signs
\end{itemize}
\end{frame}

\begin{frame}{Common Applications}
\begin{itemize}
    \item \textbf{Geometry:} Ladders, shadows, expanding shapes
    \item \textbf{Physics:} Moving objects, angles, distances
    \item \textbf{Engineering:} Tanks, pipes, mechanical systems
    \item \textbf{Real-world:} Traffic, weather, economics
\end{itemize}

\vspace{0.5cm}
Related rates problems connect calculus to real-world applications and help develop problem-solving skills.
\end{frame}

\begin{frame}{Thank You!}
\centering
\vspace{2cm}
{\Huge \textcolor{myblue}{\textbf{Questions?}}}

\vspace{1cm}
{\Large Related rates show how calculus applies to the real world!}
\end{frame}

\end{document} 